---
title: "Field et al. Model"
author: "Libby Pansing"
date: "10/17/2016"
output:
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
  html_document: default
---
```{r, message = F, echo = F, warning=FALSE}
# Field et al. model development

library(popbio)
library(dplyr)
library(magrittr)
library(pander)

# test pop size

n <- c(62+353, 580, 38, 79, 65, 91)
```

The following code implements the population model developed by Field et al. (2012). In this implementation, we consider their model without any disease components. 


Field et al. estimates population size at time = t given a stage structured population with size $\vec{x}$. They consider six stages: 1) seed (SEEDS), 2) primary seedling (SD1), 3) secondary seedling (SD2), 4) sapling (SA), 5) young adult (YA), and 6) mature adult (MA). These class designations are based on diameter at breast height (dbh). 

Their estimates rely on the following sequence of events:

Leaf area index (LAI) of population size at time = t -> seedling recruitment ($\vec{\gamma}$) -> survival and transition, which creates an intermediate population size y -> LAI of intermediate population size y -> fecundity -> pop size t = t+1


Field et al. used the following non-linear map to estimate class structured population size at time = t + 1:

$$\vec{x}_{6}(t + 1) = \vec{g}_6[\vec{x}_6(t)] + \vec{f}_6(\vec{g}_6[\vec{x}_6(t)])$$

where $\vec{g}_6$ and $\vec{f}_6$ are vector valued functions that estimate survivorship and transition, and fecundity, respecitvely.


$\vec{g}_6$ is divided into linear and non linear components:

$$ \vec{g}_6(\vec{x}_6) = \boldsymbol{S} * (\vec{x}_6 + \vec{\gamma}_6(\vec{x}_6)) $$
where **S** is the survival and transition matrix comprising the linear components, and $\vec{\gamma}(\vec{x})$

$$\vec{\bar{y_6}} = \vec{x_6} + \vec{\gamma_6} (\vec{x_6})$$
I begin by defining the variables necessary to estimate $\vec{\gamma_6}(\vec{x_6})$, which estimates the germination probability, or the probability of transition from seed to primary seedling.




##Define variables

**Parameters to estimate life-stage specific anual survival**
```{r}
#define variables

mortality.seeds  <- function(){  # mortality rate of seeds
  return(1)
}


mortality.SD1    <- function(){  # mortality rate of primary seedlings
  return(0.152)
}


mortality.SD2    <- function(){  # mortality rate of secondary seedlings 
  return(0.105)
}


mortality.SA     <- function(){  # mortality rate of saplings
  return(0.020)
}


mortality.YA     <- function(){  # mortality rate of young adults
  return(0.015)
}

mortality.MA     <- function(){  # mortality rate of mature adults
  return(0.05)
}


mortality.vector <- function(x){   #mortality vector
  c(mortality.seeds(),
    mortality.SD1(),
    mortality.SD2(),
    mortality.SA(),
    mortality.YA(),
    mortality.MA())
}

mortality.vector()

# Residence time
residence.seeds  <- function(x){   # # of years as seeds
  return(1) 
}


residence.SD1 <- function(x){   # # years as primary seedling (SD1)
  return(4) 
}

residence.SD2  <- function(x){   # # years as 2ndary seedling (SD2)
  return(16) 
}


residence.SA   <- function(x){   # # years as sapling (SAP)
  return(20) 
}

residence.YA   <- function(x){   # # years as young adult (YA)
  return(50) 
}   

residence.MA   <- function(x){
  return(Inf)
}
                     


residence.vector <- function(x){
  c(residence.seeds(),
    residence.SD1(),
    residence.SD2(),
    residence.SA(),
    residence.YA(),
    residence.MA())
}

residence.vector()

```

**Survial estimates**
```{r}
si <- function(){
  (1 - (1/residence.vector())) *
  (1 - mortality.vector())
}

  
ti <- function() {
  (1/residence.vector()) * 
  (1 - mortality.vector())
}


S <- function(){
  matrix(c(si()[1],  0,  0,  0,  0,  0,
              ti()[1], si()[2],  0,  0,  0,  0,
              0, ti()[2], si()[3],  0,  0,  0,
              0,  0, ti()[3], si()[4],  0,  0, 
              0,  0,  0, ti()[4], si()[5],  0,
              0,  0,  0,  0, ti()[5], si()[6]),
            byrow = T, nrow = 6) 
}

S()
```

**Parameters to estimate LAI**
```{r}
#mean dbh for each stage
d1 <- function() {   # dbh seeds
  return(0)
}

d2 <- function() {   #dbh SD1 
  return(0)
}

d3 <- function(){ #dbh SD2
  return(0)
}

d4 <- function(){ #dbh SA
  return(2.05)
}

d5 <- function(){ # dbh YA
  return(12.5)
}

d6 <-function() { #dbh MA
  return(37.0)
}

## Leaf area coefficients. Define the relationship between leaf area and diameter. Estimated via MLE assuming the general form y = ax^b

alpha1 <- function(){
  return(0.456)
}

alpha2 <- function(){
  return(0.0736)
}

alpha3 <- function(){
  return(2.070)
}


l <- function(){
  c(d1(), 
    d2(), 
    alpha1(), 
    alpha2() * d4() ^ alpha3(), 
    alpha2() * d5() ^ alpha3(),
    alpha2() * d6() ^ alpha3())
}


LAIb    <- function(){     # Background leaf area index
  return(0)
}


LAI <- function(x) {       # LAI of the study area
  l <- l()
  return((t(l) %*% x)/10000 + LAIb())
}

LAI(n)

yhat <- function(x){  # function that estimates intermediate population size y
  y <- S() %*% x
  return(y)
}

yhat(n)
```

**Fecundity parameters**
```{r}
Cmax    <- function(){     # Maximum number of cones per tree
  return(7.5)
}

Cmax()

Scone   <- function(){     # Number of seeds per cone
  return(46)
}

Scone()

nBirds  <- function(){     # Number of Clark's nutcrackers in the area
  return(3)
}

nBirds()


Pfind    <- function(){    # Probability of finding seeds
  return(0.8)
} 

Pfind()

Pcons    <- function(){    # Probability of consuming seeds
  return(0.3)
}

Pcons()

SpC      <- function() {   # Number of seeds per cache
  return(3.7)
}

SpC()


FRatio   <- function(){    # Fecundity ratio, # of seeds per
  return(0.1)
}

FRatio()

```


Nonlinear components

```{r nonlinear components}
SpB <- function(x) {
  x[1]/nBirds()
  }

SpB(n)
# Number of seeds cached per bird, where
                # x = number of seeds at time t and                     # nBirds = nutcracker density                           # (birds/hectare)

rcache <- function(x){
  0.73/(1 + exp((31000 - SpB(x))/3000))
              # reduction factor for the propensity to 
              # cache seeds
} 

rcache(n)

rALs <-   function(x){
  1/(1 + exp(2*(LAI(x) - 3)))
              # reduction factor for the available light
              # completely dependent on LAI
} 

rALs(n)

r2        <-      function(x) {        # probability of seeds germinating in a given year
  (((1-Pfind()) * (1- Pcons()))/SpC()) * rcache(x) * rALs(x)
              
}

r2(n)

Ctree <-  function(x){            # number of cones per tree
   ((0.5/(1 + exp(5*(LAI(x)) - 2.25)))+0.5) * Cmax() 
} 
              

Ctree(n)

r1 <- function(x){
  Scone() * Ctree(x)
}

r1(n)


rho <- 0.1 #
e1  <- c(1, 0, 0, 0, 0, 0)
e2  <- c(0, 1, 0, 0, 0, 0)

f <- function(y){
  f <- r1(y) * (rho * y[5] + y[6]) * e1
  return(f)
}

```


```{r}

tplus1 <- function(x){
  LAIt <- LAI(x)  # Leaf area index of the initial pop size
  t2   <- r2(x) * x[1] * e2 # probab of germination:t2 (transition prob SD -> SD1)
  yt   <- S() %*% (x + t2)  # incorporate t2 into the transition matrix & calculate intermediate population size y
  LAIy <- LAI(yt) # Leaf area index of intermediate population, y. 
  ft   <- f(yt)   # Fecundity based on pop size y and LAI(y)
  nt1  <- yt + ft # Add cone production to the pop size y to obtain n(t + 1)
  return(nt1)
}

n0 <- n
popsize <- NULL
th <- 100
popsize <- matrix(rep(0, length(n) * th), nrow = length(n))

# function body is this....
#args n and th
for(i in 1:th){
  popsize[,i] <- tplus1(n)
  n <- tplus1(n) #add stochastic components here 
}


# but this can be vectorized and will result in an array (stacks of outcomes)






wbp.projection <- function(popSize = mat, Years = y){
  if (ncol(mat == Years)){
    return(nt1)
  } else
  n    <- x
  LAIt <- LAI(n)  # Leaf area index of the initial pop size
  t2   <- r2(n) * n[1] * e2 # probab of germination:t2 (transition prob SD -> SD1)
  yt   <- S() %*% (n + t2)  # incorporate t2 into the transition matrix & calculate intermediate population size y
  LAIy <- LAI(yt) # Leaf area index of intermediate population, y. 
  ft   <- f(yt)   # Fecundity based on pop size y and LAI(y)
  nt1  <- yt + ft # Add cone production to the pop size y to obtain n(t + 1)
  if(length(nt1 == 6)){
    nt1 <- append(n, nt1, after = length(n))
  } else{
    nt1 <- append(nt1, testThisShit(nt1), after = length(nt1))
  }
}

wbp.projection(n, Years = 5)



r <- function(mat, it){
  if(ncol(mat) < it){
    mat2 <- matrix(c(1:6), ncol = 1)
    cbind(mat,mat2)
    mat <- r(cbind(mat,mat2), it)
  }
  return(mat)
}

r(test, 5)

n <- matrix(c(16000, 580,  38,  79,  65,  91), nrow = 6)
popSize <- n


wbp.projection <- function(popSize = mat, Years = y){
  if (ncol(popSize) < Years){
  LAIt <- LAI(popSize[,ncol(popSize)])  # Leaf area index of the initial pop size
  t2   <- r2(popSize[,ncol(popSize)]) * popSize[2, ncol(popSize)] * e2 # probab of germination:t2 (transition prob SD -> SD1)
  yt   <- S() %*% (popSize[,ncol(popSize)] + t2)  # incorporate t2 into the transition matrix & calculate intermediate population size y
  LAIy <- LAI(yt) # Leaf area index of intermediate population, y. 
  ft   <- f(yt)   # Fecundity based on pop size y and LAI(y)
  nt1  <- matrix(yt + ft, ncol = 1) # Add cone production to the pop size y to obtain n(t + 1)
  popSize <- testThisShit(cbind(popSize, nt1), Years)
  }
  return(popSize)
}


test <- wbp.projection(popSize = n, Years = 20)
plot(colSums(test))
```
